---
title: "2 Numeric Variables Guided Lab 2025"
author: "Robert T Kohn"
date: "9/2/25"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

# Importing & Cleaning Data

```{r}
library(readr)
Quarterbacks <- readr::read_csv("C:/Users/RobTop/Documents/STA227 - Data Visualization/Quarterbacks.csv")
Quarterbacks<-na.omit(Quarterbacks)#Remove 2 Rows With Errors or Missing value
head(Quarterbacks)

```
# Exploring Data

I would say theres probably a decent argument for those who play more have a higher score however players like Barry Sanders would likely have a similar score to some 'one hit wonder' players due to lack of games in comparison to a few big season players. Curious to see what happens.

# Application
```{r}
gf_point( HOFScore ~ GamesStarted , data = Quarterbacks, title="Scatterplot Starts vs HOFScore")
gf_point( GamesStarted ~ HOFScore , data = Quarterbacks, title="Scatterplot HOFScore vs Starts")
gf_point( GamesStarted ~ HOFScore , data = Quarterbacks , alpha = 0.5 , title="Scatterplot HOFScore vs Starts , alpha = 0.5 ")
```

# Contour Lines & Layers
```{r}
gf_density_2d( HOFScore ~ GamesStarted , data = Quarterbacks, title="gf_density_2d Starts vs HOFScore")
gf_density_2d( HOFScore ~ GamesStarted , data = Quarterbacks, title="contour lines and each individual point") %>% gf_point(HOFScore ~ GamesStarted , data = Quarterbacks , alpha = 0.5 )
```

# Density Filled (With Layers)
```{r}
gf_density_2d_filled( HOFScore ~ GamesStarted , data = Quarterbacks, title="gf_density_2d Starts vs HOFScore Filled")
gf_density_2d_filled( HOFScore ~ GamesStarted , data = Quarterbacks, title="contour lines and each individual point Filled") %>% gf_point(HOFScore ~ GamesStarted , data = Quarterbacks , alpha = 0.5, color="white")
```

# Hex Plot
```{r}
gf_hex( HOFScore ~ GamesStarted , data = Quarterbacks, title="Hexplot Starts vs HOFScore")
```
