---
title: "Categorical Data Guided Lab"
author: "Robert T Kohn"
date: "9/2/25"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```
# Load Athritis Package

```{r}

require(vcd)
data("Arthritis")
head(Arthritis)

```

# Planning the exploration
It is important to have a guided question for each visual because otherwise we can form biases off the data and form a hypothesis to match the work we have done which then causes the work to be flawed. Additionally, a hypothesis is fundamental with all statistical analysis.

I don't know how to embed an image into markdown but imagine a chart similar to yours but some is reduced for treatment.

# Bar Charts

```{r}

gf_bar( ~ Treatment , data = Arthritis, title="Arthritis Bar Chart")
gf_bar( ~ Treatment , data = Arthritis , fill = ~ Improved, title="Filled in Arthritis Bar Chart" )
gf_bar( ~ Treatment , data = Arthritis , fill = ~ Improved , position = position_fill( ),title="Position Filled in Arthritis Bar Chart"  )

```

# Side Bar Charts

```{r}

gf_bar( ~ Treatment , data = Arthritis , fill = ~ Improved,position = position_dodge( ), title="Filled in Arthritis Bar Chart" )
gf_bar(~Improved, data = Arthritis) %>% gf_facet_wrap(~Treatment)
gf_bar(~Improved, data = Arthritis, fill = ~Improved, show.legend = FALSE) %>%
gf_facet_wrap(~Treatment)


```

# Alt Charts

```{r}

gf_count( Improved ~ Treatment, data = Arthritis )
gf_bin2d( Improved ~ Treatment , data = Arthritis )
gf_bin2d(Improved~Treatment, data=Arthritis) %>%
gf_refine(scale_fill_continuous(low = "white", high = "navy"))

```


# Load Hair and Eye data

```{r}

data("HairEyeColor")
HairEye <- as.data.frame(HairEyeColor)
head(HairEyeColor)

```

# Planning Exploration

I have no clue how to embed yet again but I'm going to list each in descending order such that Hair={Eye,Eye,...}. Black={Brown,Blue,Hazel,Green} Brown={Brown,Blue,Hazel,Green} Red={Brown,Blue,Hazel,Green} and Blond={Blue,Green,Hazel,Brown}.


# Column Chart

```{r}
library(ggformula)
gf_col(Freq ~ Hair, data = HairEye, fill = ~Eye)
gf_col(Freq~Eye, data =HairEye, fill = ~Hair) %>%
gf_refine(scale_fill_manual(values = c("black", "#815214", "tomato", "lightgoldenrod1")))
gf_col(Freq ~ Hair, data = HairEye, fill = ~Eye) %>% gf_refine(scale_fill_manual(values = c("#815214", "dodgerblue", "gold", "springgreen3")))
gf_col(Freq ~ Hair, data = HairEye, fill = ~Eye, position = position_fill())
gf_col(Freq ~ Hair, data = HairEye, fill = ~Eye, position = position_dodge())

```

# Count Chart

```{r}
gf_point(Freq ~ Hair, data = HairEye, fill = ~Eye)
gf_tile(Freq ~ Hair, data = HairEye, fill = ~Eye)
gf_tile(Hair~Eye, fill=~Freq, data = HairEye) %>%
gf_refine(scale_fill_continuous(low = "white", high = "navy"))


```