---
title: "Kohn Dashboard Capstone Project"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
    theme: flatly
    source_code: embed
    #runtime: shiny
    navbar:
      - { title: "Data Paste", icon: "fa-code", href: "https://pastebin.com/KRy3keEw", align: right }
---

```{r setup, include=FALSE}
library(flexdashboard)
library(DT)
library(plotly)
library(ggplot2)
library(dplyr)
library(fontawesome)
library(shiny)
library(dplyr)
library(stringr)
library(tidyverse)
library(plotly)
library(tidyr)
library(MASS)
library(mosaic)
library(ovml)
library(lubridate)
library(readr)
library(viridis)
library(fields)
set.seed(67)
```
# **Volleyball Hub**

### Welcome Header {.no-title}
```{r welcome-header, echo=FALSE}
tags$div(
  style = "background: linear-gradient(135deg, #1e3c72, #2a5298); 
           padding: 20px; border-radius: 10px; color: white; text-align: center;",
  h1("🏐 Volleyball Analytics Dashboard", style = "margin-bottom: 10px;"),
  h4("Analyzing VNL 2025 Grand Finals, PlusLiga 2008-2023, and VNL Beach 2025", 
     style = "opacity: 0.9;"),
  tags$p("Interactive Dashboard • December 2025", style = "font-style: italic;")
)
```

## **League Quick Facts**

Row {data-height=220}
-----------------------------------------------------------------------

### **PlusLiga 2024/25 🇵🇱**

```{r quick-stats-pl}
tags$div(
  style = "display: grid; grid-template-columns: repeat(3, 1fr); gap: 10px; padding: 15px;",
  
  # Season Length
  tags$div(
    style = "background: linear-gradient(135deg, #1e3c72, #2980b9); 
             padding: 15px; border-radius: 8px; color: white; text-align: center;",
    tags$h3("8 Months", style = "margin: 0; font-size: 24px;"),
    tags$p(icon("calendar"), style = "font-size: 20px; margin: 5px 0;"),
    tags$p("Oct - May", style = "margin: 0; font-size: 12px; opacity: 0.9;")
  ),
  
  # Total Teams
  tags$div(
    style = "background: linear-gradient(135deg, #2980b9, #3498db); 
             padding: 15px; border-radius: 8px; color: white; text-align: center;",
    tags$h3("16 Teams", style = "margin: 0; font-size: 24px;"),
    tags$p(icon("users"), style = "font-size: 20px; margin: 5px 0;"),
    tags$p("Professional Clubs", style = "margin: 0; font-size: 12px; opacity: 0.9;")
  ),
  
  # Prize Pool
  tags$div(
    style = "background: linear-gradient(135deg, #3498db, #85c1e9); 
             padding: 15px; border-radius: 8px; color: #2c3e50; text-align: center;
             border: 2px solid #1e3c72;",
    tags$h3("€1.5M+", style = "margin: 0; font-size: 24px; color: #1e3c72;"),
    tags$p(icon("trophy"), style = "font-size: 20px; margin: 5px 0; color: #f39c12;"),
    tags$p("Total Prize Pool", style = "margin: 0; font-size: 12px; opacity: 0.8;")
  )
)
```

### **VNL Indoor Facts 🏐**

```{r quick-stats-vnl}
tags$div(
  style = "display: grid; grid-template-columns: repeat(4, 1fr); gap: 10px; padding: 15px;",
  
  # Tournament Duration
  tags$div(
    style = "background: linear-gradient(135deg, #27ae60, #2ecc71); 
             padding: 15px; border-radius: 8px; color: white; text-align: center;",
    tags$h3("6 Weeks", style = "margin: 0; font-size: 24px;"),
    tags$p(icon("clock"), style = "font-size: 20px; margin: 5px 0;"),
    tags$p("June - July", style = "margin: 0; font-size: 12px; opacity: 0.9;")
  ),
  
  # Total Teams
  tags$div(
    style = "background: linear-gradient(135deg, #2ecc71, #58d68d); 
             padding: 15px; border-radius: 8px; color: white; text-align: center;",
    tags$h3("24 Teams", style = "margin: 0; font-size: 24px;"),
    tags$p(icon("users"), style = "font-size: 20px; margin: 5px 0;"),
    tags$p("18 Nations", style = "margin: 0; font-size: 12px; opacity: 0.9;")
  ),
  
  # Global Reach
  tags$div(
    style = "background: linear-gradient(135deg, #58d68d, #abebc6); 
             padding: 15px; border-radius: 8px; color: #2c3e50; text-align: center;",
    tags$h3("16+ Nations", style = "margin: 0; font-size: 24px;"),
    tags$p(icon("globe"), style = "font-size: 20px; margin: 5px 0; color: #27ae60;"),
    tags$p("Participating Countries", style = "margin: 0; font-size: 12px; opacity: 0.8;")
  ),
  
  # Prize Pool
  tags$div(
    style = "background: linear-gradient(135deg, #abebc6, #e8f8f5); 
             padding: 15px; border-radius: 8px; color: #2c3e50; text-align: center; 
             border: 2px solid #27ae60;",
    tags$h3("$4M+", style = "margin: 0; font-size: 24px; color: #27ae60;"),
    tags$p(icon("trophy"), style = "font-size: 20px; margin: 5px 0; color: #e67e22;"),
    tags$p("Total Prize Pool", style = "margin: 0; font-size: 12px; opacity: 0.8;")
  )
)
```

### **Beach Pro Tour Facts 🏖️**

```{r quick-stats-bpt}
tags$div(
  style = "display: grid; grid-template-columns: repeat(3, 1fr); gap: 10px; padding: 15px;",
  
  # Global Stops
  tags$div(
    style = "background: linear-gradient(135deg, #e67e22, #f39c12); 
             padding: 15px; border-radius: 8px; color: white; text-align: center;",
    tags$h3("14+ Stops", style = "margin: 0; font-size: 24px;"),
    tags$p(icon("map-marker-alt"), style = "font-size: 20px; margin: 5px 0;"),
    tags$p("Global Locations", style = "margin: 0; font-size: 12px; opacity: 0.9;")
  ),
  
  # Continents
  tags$div(
    style = "background: linear-gradient(135deg, #f39c12, #f8c471); 
             padding: 15px; border-radius: 8px; color: white; text-align: center;",
    tags$h3("6 Continents", style = "margin: 0; font-size: 24px;"),
    tags$p(icon("globe-americas"), style = "font-size: 20px; margin: 5px 0;"),
    tags$p("Worldwide Reach", style = "margin: 0; font-size: 12px; opacity: 0.9;")
  ),
  
  # Elite Events
  tags$div(
    style = "background: linear-gradient(135deg, #f8c471, #fdebd0); 
             padding: 15px; border-radius: 8px; color: #2c3e50; text-align: center;
             border: 2px solid #e67e22;",
    tags$h3("Elite 16", style = "margin: 0; font-size: 24px; color: #e67e22;"),
    tags$p(icon("crown"), style = "font-size: 20px; margin: 5px 0; color: #f1c40f;"),
    tags$p("Top Tier Events", style = "margin: 0; font-size: 12px; opacity: 0.8;")
  )
)
```

Row {data-height=200}
-----------------------------------------------------------------------

### **🌍 Geographic Scope**
```{r}

tags$div(
  style = "padding: 20px; background-color: #e8f4f8; border-radius: 8px;",
  h4("The Leagues:"),
  tags$ul(
    tags$li(tags$strong("PlusLiga:"), " National league (Poland only)"),
    tags$li(tags$strong("VNL:"), "There are 18 nations who compete in the FIVB VNL. These countries are multi-continental with teams such as China, Cuba, Bulgaria, Turkey, and many more."),
    tags$li(tags$strong("Beach Tour:"), " Global circuit (40+ countries)")
  )
)

```

### **Geographic Reach**
```{r}

valueBox("Year-Round", "Continuous Competition", 
         icon = fa_i("calendar-alt"), 
         color = "info")

```

# **Overview** {#overview}

Row {data-height=450}
-----------------------------------------------------------------------

### **Why This Matters**
```{r purpose}
tags$div(
  style = "padding: 20px; background: linear-gradient(135deg, #f8f9fa, #e9ecef); 
           border-radius: 10px;",
  h3("🏐 Why Study Volleyball Analytics?"),
  p("Volleyball represents one of the fastest-growing professional sports globally, yet its statistical analysis lags behind sports like basketball and soccer. These visualizations aim to bridge that gap by:"),
  tags$ul(
    tags$li("📈 Identifying key performance indicators that predict team success"),
    tags$li("🔄 Analyzing how age can coorelate to success"),
    tags$li("🏐 Comparing where the ball travels over the course of the game"),
    tags$li("🔮 Examining Common Trends")
  ),
  br(),
  p("As a data enthusiast and volleyball fan, I believe data-driven insights can enhance both team performance and fan engagement in the sport.")
)

```


Row {data-height=200}
-----------------------------------------------------------------------

```{r}

tags$div(
  style = "padding: 20px; background-color: #e8f4f8; border-radius: 10px;",
  h4("🎯 Personal Motivation"),
  p("My interest in volleyball analytics stems from:"),
  tags$ul(
    tags$li("Playing intramural volleyball for 2 years"),
    tags$li("Observing how data transformed other professional sports"),
    tags$li("Noticing limited statistical resources available to volleyball enthusiasts"),
    tags$li("Wanting to dip my feet into sport analytics")
  )
)

```


```{r}

tags$div(
  style = "padding: 20px; background-color: #f3e8ff; border-radius: 10px;",
  h4("🎯 Key Questions"),
  p("This dashboard seeks to answer:"),
  tags$ul(
    tags$li("Is there a difference in duration between genders?"),
    tags$li("Does age effect winning probability?"),
    tags$li("What separates elite beach volleyball pairs from average ones?"),
    tags$li("How impactful is the middle of the court?")
  )
)

```


# **VNL 2025 Finals** {#VNL}

## **Data Analysis Sections**

Row {data-height=450}
-----------------------------------------------------------------------

### VNL 2025 Grand Finals {data-icon="fa-trophy"}

```{r vnl-section}
ball_data <- read_csv("C:/Users/RobTop/Documents/STA227 - Data Visualization/vb/ball_positions.csv")
ball_data$x_court <- -9 + ball_data$u * 18  # u in [0,1] mapped to [-9,9]
y_min_clip <- 2.9
y_max_clip <- 9
ball_data$y_court <- pmax(pmin(ball_data$v * 9, y_max_clip), y_min_clip)  # clip
ball_data$y_court <- (ball_data$y_court - y_min_clip) / (y_max_clip - y_min_clip) * 9  # stretch to 0-9

z <- kde2d(
  x = ball_data$x_court,
  y = ball_data$y_court,
  n = 200,
  lims = c(-9, 9, 0, 9)
)
z$z <- sqrt(sqrt(z$z))  # more sensitive

# Enhanced plot setup
par(mar = c(4, 3.5, 4, 8))

# Create enhanced background with more right space
plot.new()
plot.window(xlim = c(-10, 12.5), ylim = c(-1, 10), xaxs = "i", yaxs = "i")
rect(-10, -1, 12.5, 10, col = "gray95", border = NA)

# Plot heatmap with enhanced colors
image(
  z,
  col = viridis(30, option = "plasma", alpha = 0.55),
  xlim = c(-9, 9),
  ylim = c(0, 9),
  axes = FALSE,
  xlab = "",
  ylab = "",
  add = TRUE
)

# Enhanced title
title(main = "Volleyball Ball Position Heatmap", 
      line = 2.5, cex.main = 1.4, font.main = 2, col.main = "gray20")
mtext("Court Position Density During Play", 
      line = 1, cex = 1, col = "gray40")

# Enhanced court boundaries
rect(-9, 0, 9, 9, border = "gray30", lwd = 2.5)
abline(v = 0, lwd = 2.5, col = "gray30")
abline(v = -3, lty = "dashed", lwd = 1.5, col = "gray50")
abline(v = 3, lty = "dashed", lwd = 1.5, col = "gray50")

# Enhanced court shading
rect(-9, 0, 0, 9, border = NA, col = rgb(0.2, 0.4, 0.8, 0.08))
rect(0, 0, 9, 9, border = NA, col = rgb(0.8, 0.2, 0.2, 0.08))
rect(-9, 0, -3, 9, border = NA, col = rgb(0.2, 0.4, 0.8, 0.12))
rect(-3, 0, 0, 9, border = NA, col = rgb(0.2, 0.4, 0.8, 0.05))
rect(3, 0, 9, 9, border = NA, col = rgb(0.8, 0.2, 0.2, 0.12))
rect(0, 0, 3, 9, border = NA, col = rgb(0.8, 0.2, 0.2, 0.05))

# Axis labels
mtext("Court Length (meters)", side = 1, line = 2.2, cex = 1.1, col = "gray30")
mtext("Court Width (meters)", side = 2, line = 2.2, cex = 1.1, col = "gray30")

# Reference grid
abline(h = seq(0, 9, by = 3), lty = "dotted", col = "gray80", lwd = 0.7)
abline(v = seq(-9, 9, by = 3), lty = "dotted", col = "gray80", lwd = 0.7)

# Enhanced team labels
text(-4.5, 8.5, "POLAND", cex = 1.1, font = 2, col = "navyblue")
text(4.5, 8.5, "ITALY", cex = 1.1, font = 2, col = "firebrick")

# Zone labels
text(-6, 3.5, "Defense Zone", cex = 0.8, col = "#50C878", srt = 90)
text(-1.5, 3.5, "Attack Zone", cex = 0.8, col = "#50C878", srt = 90)
text(1.5, 3.5, "Attack Zone", cex = 0.8, col = "#50C878", srt = 90)
text(6, 3.5, "Defense Zone", cex = 0.8, col = "#50C878", srt = 90)

# Enhanced legend
cols <- viridis(30, option = "plasma", alpha = 0.85)
legend_x <- 9.1  
legend_y <- seq(0, 9, length.out = length(cols) + 1)

# Legend background
rect(legend_x - 0.1, -0.2, legend_x + 1.3, 9.2, col = "white", border = "gray70", lwd = 0.8)

# Color gradient
for(i in 1:length(cols)){
  rect(legend_x, legend_y[i], legend_x + 0.5, legend_y[i+1], 
       col = cols[i], border = NA)
}

# Legend title
text(legend_x + 0.25, 9.3, "Ball Density", adj = 0.5, cex = 1, font = 2, col = "gray30")

# Numerical labels
text(legend_x + 0.6, seq(0, 9, length.out = 5), 
     labels = sprintf("%.2f", seq(min(z$z), max(z$z), length.out = 5)), 
     adj = 0, cex = 0.85, col = "gray40")

# Net indicator
segments(-0.1, 0, -0.1, 9, lwd = 3, col = "gray50")
segments(0.1, 0, 0.1, 9, lwd = 3, col = "gray50")
text(0, -0.7, "NET", cex = 0.9, font = 2, col = "gray40")  

# Data caption
#mtext(paste("Data points: n =", nrow(ball_data)), side = 1, line = 3.5, cex = 0.8, col = "gray50", adj = 0)

hist(ball_data$y_court, breaks = 20, main = "Distribution of Ball Positions Along Court Y-Axis", xlab = "Court Y Position", ylab = "Frequency", col = "steelblue", border = "white", cex.main = 1.3, cex.lab = 1.1, cex.axis = 1, font.main = 2, las = 1, tcl = -0.3, mgp = c(2.5, 0.7, 0), bty = "n")

tags$div(
  style = "padding: 20px;",
  h3("🏆 VNL 2025 Grand Finals"),
  p("This analysis used the Italy vs Poland VNL 2025 match video, tracking ball",
    " movement throughout the game to identify spatial patterns and tactical",
    " tendencies using computer vision and the OpenVolley package."),
  br(),
  p("Volleyball is an intense sport where the ball travels all over the court but there has to be a pattern to maintain the flow without it being pure chaos. This is clearly illustrated by both the heatmap and the histogram. Starting with the histogram, the middle of the court is clearly the most intense place and there is a lot of reasons for this but there are a lot of common reasons. Service takes place from the center back and so while they wait for the whistle they hold the ball and this is footage of the whole game. Additionally, the setter operates as a central hub at the net. Their positioning maximizes offensive options and includes the threat of a dump which further concentrates events in this region. Furthermore, a high volume of passes, sets, and quick attacks are funneled through this central corridor. Consequently, with setters, the Middle Blocker can be used as a decoy or the greatest weapon. The middle is a battle of height and who can reach the peak of their jump first or who can get the highest. Either way, it is an intense place which draws attention. For example, a successful middle-blocker decoy often creates isolated blocking opportunities for the outside hitters attacking from the pins. All of these factors can contribute to why the middle of the court sees so much play. The heatmap, conversely, shows clear density spikes at the court's edges and the attack line. This makes it very clear where setters will send the ball. These spikes correspond directly to the primary attack zones of specific position players. Typically the edges belong to the Outside or Opposite Hitters. These two are responsbile for fighting at the pins and will either hit crosses to go across the court and negate the blocks or try and hit a line shot to get on the edge of the court. The density clusters just behind the attack line are often the result of strategic back-row attacks. When a primary hitter rotates to the back row, setters frequently utilize them to keep the offense multi-dimensional and pressure the defense. This pattern indicates a deliberate tactical preference to leverage powerful back-row attackers, like the opposite hitter, as a consistent offensive option from deep positions. Secondary density marks throughout the court typically indicate successful tactical shots, such as sharp-angle attacks or tips into defensive gaps created by the opponent's formation. In professional volleyball the gaps close quickly due to blockers and typically due to the formation of such blocks the cross shot typically lands in the center and if one is a middle blocker they usually want to try and bounce the ball or waterfall off their blocker."),
  br(),
  p("All in all, this heatmap provided fascinating insight and it was a very interesting process to make work.")
)
```



## **Data Source Information**

Row {data-height=300}
-----------------------------------------------------------------------

### About the Data {#datasource}

```{r data-source}
tags$div(
  style = "padding: 20px;",
  h3("📋 Data Sources"),
  tags$ul(
    tags$li(tags$strong("VNL 2025 Grand Finals:"), 
            " Ball tracking analysis from Italy vs Poland match using ",
            tags$a(href="https://www.youtube.com/watch?v=VL4WutaWt9U", 
                   "YouTube Game Recording", target="_blank"),
            " processed with ",
            tags$a(href="https://openvolley.org/", 
                   "OpenVolley/ovml R package", target="_blank"))
  ),
  hr(),
  h4("Research Question:"),
  p("Where is the ball typically located on the court over the course of a game?",
    br(),
    ),
  hr(),
  p("Dashboard created with R flexdashboard • Updated: December 2025"),

hr(),
  h4("🎬 Match Highlights:"),
  tags$div(
    style = "text-align: center; padding: 20px; background-color: #f8f9fa; border-radius: 8px;",
    
    # Clickable YouTube thumbnail
    tags$a(
      href = "https://youtu.be/owC7X8hTin4?t=39",
      target = "_blank",
      style = "display: inline-block; text-decoration: none;",
      tags$div(
        style = "position: relative; width: 100%; max-width: 560px; margin: 0 auto;",
        
        # Thumbnail image
        tags$img(
          src = "https://img.youtube.com/vi/owC7X8hTin4/hqdefault.jpg",
          style = "width: 100%; border-radius: 8px; box-shadow: 0 4px 8px rgba(0,0,0,0.2);",
          alt = "Italy vs Poland VNL 2025 Highlights"
        ),
        
        # Play button overlay
        tags$div(
          style = "position: absolute; top: 50%; left: 50%; transform: translate(-50%, -50%); 
                   background-color: rgba(255, 0, 0, 0.9); width: 70px; height: 50px; 
                   border-radius: 10px; display: flex; align-items: center; justify-content: center;",
          icon("play", style = "color: white; font-size: 24px; margin-left: 5px;")
        ),
        
        # Video title
        tags$div(
          style = "position: absolute; bottom: 0; left: 0; right: 0; 
                   background: linear-gradient(transparent, rgba(0,0,0,0.7)); 
                   color: white; padding: 10px; border-bottom-left-radius: 8px; 
                   border-bottom-right-radius: 8px; text-align: left;",
          tags$p(
            style = "margin: 0; font-weight: bold;",
              "Italy vs Poland VNL 2025 Highlights"
          ),
          tags$p(
            style = "margin: 0; font-size: 12px; opacity: 0.9;",
            "Click to watch (starts at 0:39)"
          )
        )
      )
    ),
    
    br(),
    
    # Direct link as backup
    tags$div(
      style = "margin-top: 15px;",
      tags$a(
        href = "https://youtu.be/owC7X8hTin4?t=39",
        target = "_blank",
        style = "color: #ff0000; text-decoration: none; font-weight: bold;",
        icon("external-link-alt"),
        " Open highlights in YouTube"
      )
    )
  )
)
```

# **VNL Beach 2025** {#beach}

## **Data Analysis Sections**

Row {data-height=450}
-----------------------------------------------------------------------


### VNL Beach 2025 {data-icon="fa-umbrella-beach"}

```{r beach-section}
vb_data  <- read_csv("C:/Users/RobTop/Documents/STA227 - Data Visualization/vb_matches.csv")
vb_clean <- vb_data %>% mutate(duration_min = as.numeric(hms(duration)) / 60,w_p1_age = as.numeric(w_p1_age), l_p1_age = as.numeric(l_p1_age))
#head(vb_data)

# Graph 1: Match duration by gender 
gf_histogram(~ duration_min, fill = ~ gender, data = vb_clean, bins = 20, alpha = 0.6) %>% gf_facet_wrap(~ gender, nrow = 2) %>%gf_labs(title = "Male Beach Matches Tend to be Longer", subtitle = "Comparison of match length patterns between male and female players", x = "Duration (minutes)", y = "Count",fill = "Gender", caption = "Data source: Kaggle: jessemostipak")%>%gf_theme( legend.position = "none", plot.title = element_text(face = "bold", size = 14, hjust = 0.5),plot.subtitle = element_text(size = 11, color = "gray40", hjust = 0.5), plot.caption = element_text(size = 9, color = "gray60", hjust = 1), strip.text = element_text(face = "bold", size = 11),  strip.background = element_rect(fill = "gray95", color = NA), panel.grid.minor = element_blank(), panel.grid.major = element_line(color = "gray90", linewidth = 0.3), axis.title = element_text(face = "bold", size = 11), axis.text = element_text(size = 10))%>%gf_refine(scale_fill_discrete(labels = c("Male", "Female")))

# Graph 2: Age distribution
age_data <- data.frame(age = c(vb_clean$w_p1_age, vb_clean$l_p1_age))
gf_histogram(~ age, data = age_data, bins = 20, fill="darkred") %>%gf_labs(title="Most players are around 30 years old",subtitle = "Age Distribution of Players",x = "Age (years)", y = "Count", caption = "Data source: Kaggle: jessemostipak")%>%gf_theme(plot.title = element_text(face = "bold", size = 14, hjust = 0.5, margin = margin(b = 8)), plot.subtitle = element_text(size = 11, color = "gray50", hjust = 0.5, margin = margin(b = 12)),plot.caption = element_text(size = 9, color = "gray60", hjust = 1, margin = margin(t = 8)),axis.title = element_text(face = "bold", size = 11),axis.title.y = element_text(margin = margin(r = 10)),axis.title.x = element_text(margin = margin(t = 10)), axis.text = element_text(size = 10, color = "black"),axis.line = element_line(color = "black", linewidth = 0.5),axis.ticks = element_line(color = "black", linewidth = 0.5),panel.grid.major.y = element_line(color = "gray90", linewidth = 0.3),  panel.grid.minor = element_blank(), plot.margin = margin(15, 15, 15, 15))

# Graph 3: Age comparison winners vs losers
all_ages <- data.frame(age = c(vb_clean$w_p1_age, vb_clean$l_p1_age),outcome = rep(c("Winner", "Loser"), each = nrow(vb_clean)))
gf_density(~age, fill = ~outcome, data = all_ages, alpha = 0.6) %>% gf_labs(title = "Older Players tend to Win their Matches More", subtitle = "Density comparison between winners and losers", x = "Age (years)", y = "Density", fill = "Match Outcome", caption = "Data source: Kaggle: jessemostipak") %>% gf_refine(scale_x_continuous(breaks = scales::pretty_breaks(n = 8)), scale_y_continuous(expand = expansion(mult = c(0, 0.05))), theme(plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 8)), plot.subtitle = element_text(size = 12, color = "gray40", hjust = 0.5, margin = margin(b = 12)), plot.caption = element_text(size = 10, color = "gray60", hjust = 1, margin = margin(t = 10)), axis.title = element_text(face = "bold", size = 12), axis.title.y = element_text(margin = margin(r = 10)), axis.title.x = element_text(margin = margin(t = 10)), axis.text = element_text(size = 11, color = "black"), axis.line = element_line(color = "black", linewidth = 0.5), panel.grid.major = element_line(color = "gray90", linewidth = 0.3), panel.grid.minor = element_blank(), legend.position = "top", legend.title = element_text(face = "bold", size = 11), legend.text = element_text(size = 10), legend.key.size = unit(0.8, "cm"), plot.margin = margin(15, 20, 15, 15)))


# Graph 4: Age Difference

vb_clean = vb_clean %>%mutate(age_diff = w_p1_age + w_p2_age - l_p1_age - l_p2_age)
gf_histogram(~age_diff, data = vb_clean, bins = 20, fill = "steelblue", color = "white", linewidth = 0.3, alpha = 0.8) %>% gf_labs(title = "Older Players tend to be more Experienced", subtitle = "Winner's age minus loser's age", x = "Age Difference (years)", y = "Frequency", caption = "Data source: Kaggle: jessemostipak") %>% gf_refine(scale_x_continuous(breaks = scales::pretty_breaks(n = 8)), scale_y_continuous(expand = expansion(mult = c(0, 0.05))), theme(plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 8)), plot.subtitle = element_text(size = 12, color = "gray50", hjust = 0.5, margin = margin(b = 12)), plot.caption = element_text(size = 10, color = "gray60", hjust = 1, margin = margin(t = 10)), axis.title = element_text(face = "bold", size = 12), axis.title.y = element_text(margin = margin(r = 10)), axis.title.x = element_text(margin = margin(t = 10)), axis.text = element_text(size = 11, color = "black"), axis.line = element_line(color = "black", linewidth = 0.6), axis.ticks = element_line(color = "black", linewidth = 0.5), panel.grid.major.y = element_line(color = "gray90", linewidth = 0.3), panel.grid.minor = element_blank(), plot.margin = margin(15, 20, 15, 15), plot.background = element_rect(fill = "white", color = NA)), geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5, alpha = 0.7))


tags$div(
  style = "padding: 20px;",
  h3("🏖️ VNL Beach 2025"),
  p("An analysis of match duration by gender reveals distinct patterns in game flow and variability. First, we notice that the mode length of a match is around 40 minutes. However, the distribution of men's match times shows a wider spread in its central portion. There seems to be greater variablity as many games continue past 60 minutes but there are several short games too. In contrast, the women's data exhibits a tighter concentration around the mode, suggesting more consistent match durations and a more predictable game flow. The right tail for womens game is also significantly smaller and suggests male games can be drawn out more. This disparity in variability could be linked to differences in serve-pass dynamics, rally structures, or scoring efficiency between the genders."),
   br(),
  p("When exploring the ages of players there is a pronounced peak within the 27-30 year range. This suggests and idea uncommon in most sports where a players prime is in their late 20s. The distribution closely approximates a bell curve, reflecting the natural attrition of athletes at both ends of the age spectrum. It makes sense that there are very few players younger than 18 and there is a clear taper at the end from 40 to 50 or so with few examples here and there. The sustained presence of players into their 30s and 40s underscores the sport's potential for longevity at the professional level. The comparative analysis of winners' and losers' ages reveals no significant systemic age advantage. In fact one could argue that as you get older your odds to win go up. This trend is intuitively explained by the accumulation of strategic experience and technical mastery, which can offset gradual declines in physical peak. A rookie won't see the court the same way a seasoned vet would. This finding reinforces that elite volleyball rewards experience, allowing players to remain competitively viable well past their physical primes.")
)

```

## **Data Source Information**

Row {data-height=300}
-----------------------------------------------------------------------

### About the Data {#datasource}

```{r data-src}
tags$div(
  style = "padding: 20px;",
  h3("📋 Data Sources"),
  tags$ul(
    tags$li(tags$strong("VNL Beach 2025:"), 
            "Beach volleyball tournament data from Kaggle", 
            tags$a(href="https://www.kaggle.com/datasets/jessemostipak/beach-volleyball", "compiled by Jesse Mostipak",target="_blank"))
  ),
  hr(),
  h4("Research Question:"),
  p("How do non-controllable factors(Gender and Age) affect outcomes?"),
  hr(),
  p("Dashboard created with R flexdashboard • Updated: December 2025")
)
```



# **PlusLiga 2008-2023** {#liga}

## **Data Analysis Sections**

Row {data-height=450}
-----------------------------------------------------------------------

### PlusLiga 2008-2023 {data-icon="fa-chart-line"}

```{r plusliga-section}
plugliga <- read_csv("C:/Users/RobTop/Documents/STA227 - Data Visualization/Mens-Volleyball-PlusLiga-2008-2023.csv")
plugliga <- plugliga %>% mutate(Date = dmy_hm(Date), Year = year(Date), Winner = factor(Winner, levels = c(0, 1), labels = c("Team_2", "Team_1")),T1_Att_Eff = as.numeric(T1_Att_Eff), T2_Att_Eff = as.numeric(T2_Att_Eff))
#head(plugliga)

# Graph 1: Serve effectiveness comparison
gf_point(jitter(as.numeric(T2_Srv_Eff)) ~ jitter(as.numeric(T1_Srv_Eff)), data = plugliga, color = ~Winner, alpha = 0.6) %>% gf_labs(title = "Serve Efficiency doesn’t Effect who Wins", subtitle = "Jittered points with reference line of equality", x = "Team 1 Serve Efficiency (jittered)", y = "Team 2 Serve Efficiency (jittered)", caption = paste("Data: kacpergregorowicz, n =", nrow(plugliga), "matches")) %>% gf_abline(slope = 1, intercept = 0, color = "gray40", linetype = "dashed", alpha = 0.5, linewidth = 1) %>% gf_refine(theme(plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 8)), plot.subtitle = element_text(size = 12, color = "gray50", hjust = 0.5, margin = margin(b = 12)), plot.caption = element_text(size = 10, color = "gray60", hjust = 1, margin = margin(t = 10)), axis.title = element_text(face = "bold", size = 12), axis.title.y = element_text(margin = margin(r = 10)), axis.title.x = element_text(margin = margin(t = 10)), axis.text = element_text(size = 11), panel.grid.major = element_line(color = "gray90", linewidth = 0.3), panel.grid.minor = element_blank(), legend.position = "right", legend.title = element_text(face = "bold", size = 11), legend.text = element_text(size = 10), plot.margin = margin(15, 20, 15, 15)))

# Graph 2: Reception vs attack
gf_point(as.numeric(T1_Att_Kill_Perc) ~ as.numeric(T1_Rec_Perf), data = plugliga, alpha = 0.6, color = "steelblue", size = 2) %>% gf_labs(title = "Good Defense and Offense are Associated", subtitle = "Team 1 performance metrics correlation", x = "Reception Performance", y = "Attack Kill Percentage (%)", caption = paste("Data: kacpergregorowicz, n =", nrow(plugliga), "matches")) %>% gf_smooth(method = "lm", color = "firebrick", se = TRUE, alpha = 0.2, linewidth = 1) %>% gf_refine(theme(plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 8)), plot.subtitle = element_text(size = 12, color = "gray50", hjust = 0.5, margin = margin(b = 12)), plot.caption = element_text(size = 10, color = "gray60", hjust = 1, margin = margin(t = 10)), axis.title = element_text(face = "bold", size = 12), axis.title.y = element_text(margin = margin(r = 10)), axis.title.x = element_text(margin = margin(t = 10)), axis.text = element_text(size = 11), axis.line = element_line(color = "black", linewidth = 0.6), axis.ticks = element_line(color = "black", linewidth = 0.5), panel.grid.major = element_line(color = "gray90", linewidth = 0.3), plot.margin = margin(15, 20, 15, 15), plot.background = element_rect(fill = "white", color = NA)))

# Graph 3: Blocking performance Overlay
gf_histogram(~T1_Blk_Sum, data = plugliga, fill = "steelblue", alpha = 0.5, binwidth = 1, color = "white", linewidth = 0.2) %>% gf_histogram(~T2_Blk_Sum, data = plugliga, fill = "darkorange", alpha = 0.5, binwidth = 1, color = "white", linewidth = 0.2) %>% gf_labs(title = "Distribution of Blocks: Team 1 vs Team 2", subtitle = "Overlaid histograms showing block frequency distributions", x = "Number of Blocks", y = "Frequency", caption = paste("Data: kacpergregorowicz, n =", nrow(plugliga))) %>% gf_refine(theme(plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 8)), plot.subtitle = element_text(size = 12, color = "gray50", hjust = 0.5, margin = margin(b = 12)), plot.caption = element_text(size = 10, color = "gray60", hjust = 1, margin = margin(t = 10)), axis.title = element_text(face = "bold", size = 12), axis.title.y = element_text(margin = margin(r = 10)), axis.title.x = element_text(margin = margin(t = 10)), axis.text = element_text(size = 11), axis.line = element_line(color = "black", linewidth = 0.6), axis.ticks = element_line(color = "black", linewidth = 0.5), panel.grid.major.y = element_line(color = "gray90", linewidth = 0.3), panel.grid.minor = element_blank(), plot.margin = margin(15, 20, 15, 15), plot.background = element_rect(fill = "white", color = NA)), scale_fill_manual(name = "Team", values = c("steelblue" = "steelblue", "darkorange" = "darkorange"), labels = c("steelblue" = "Team 1", "darkorange" = "Team 2")), guides(fill = guide_legend(override.aes = list(alpha = 0.7))))

# Graph 4: Blocking performance Summed
gf_histogram(~(T1_Blk_Sum + T2_Blk_Sum), data = plugliga, alpha = 0.8, position = "stack", binwidth = 1, fill = "firebrick", color = "white", linewidth = 0.2) %>% gf_labs(title = "There are typically 14-19 Blocks in a Game", subtitle = "Combined blocks from both teams (T1_Blk_Sum + T2_Blk_Sum)", x = "Total Number of Blocks", y = "Frequency", caption = paste("Data: kacpergregorowicz, n =", nrow(plugliga))) %>% gf_refine(theme(plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 8)), plot.subtitle = element_text(size = 12, color = "gray50", hjust = 0.5, margin = margin(b = 12)), plot.caption = element_text(size = 10, color = "gray60", hjust = 1, margin = margin(t = 10)), axis.title = element_text(face = "bold", size = 12), axis.title.y = element_text(margin = margin(r = 10)), axis.title.x = element_text(margin = margin(t = 10)), axis.text = element_text(size = 11), axis.line = element_line(color = "black", linewidth = 0.6), axis.ticks = element_line(color = "black", linewidth = 0.5), panel.grid.major.y = element_line(color = "gray90", linewidth = 0.3), panel.grid.minor = element_blank(), plot.margin = margin(15, 20, 15, 15), plot.background = element_rect(fill = "white", color = NA)))

tags$div(
  style = "padding: 20px;",
  h3("🇵🇱️ PlusLiga"),
  p("While the data is drawn solely from Poland's PlusLiga, the league's consistent production of elite, internationally competitive players makes its data a valuable proxy for high-level volleyball trends. Surprisingly, there seems to be no real significant difference in serve effectiveness between the two teams at any given time. This implies not only a remarkable consistency in serving performance among elite players but also that match outcomes are not dictated by serving alone. This is wild to consider because the only way to get a series of points is with successful serves but this implies that what comes after the serve is more important which is poetic in a way. The idea that your team can always work together and overcome bad moments with proper team play is phenomenal and a great motivator for new players."),
   br(),
  p("Furthermore, a strong positive correlation between offensive and defensive efficiency metrics highlights the sport's integrated nature. At a very basic level this makes sense and really lays the ground work for whats important in Volleyball. If the ball never touches the ground then you can't lose the point which means you had a successful attack. But also if you're scoring points consistently while attacking its because someone got the ball to you. The correlation quantifies this chain, showing that high-performing teams are proficient in both phases, not just one. While this is a fundamental concept, the advanced interaction is where elite play is defined. If a team has shoddy receives then the setter can't get a good ball to his strong hitters which puts their team at a disadvantage and results in the other team having an easier time denying any swings regardless of how talented the setter or hitters are. Conversely, failing to terminate a high-quality attack opportunity wastes the preceding defensive effort and increases pressure on the team's own defense, creating a negative feedback loop. So not only is a basic craftsmanship level important at a basic level to play but it also allows one to become elite in the sport."),
  br(),
  p("Finally, one must consider the blocks in a game. No matter what in volleyball the opponent will score points regardless of what your team does to prevent it. Blocks are one facet of such. Knowing how likely block are to occur over the course of the game is pretty important. Knowing the importance of defense against a block and using blocking as a way of attacking is fundamental in the fast paced sport of volleyball. While this visualization does not predict the timing of blocks, it establishes a realistic expectation for their frequency and value. Unfortunately, real life is not a sports anime and every time one is at the net they cannot expect a block. However, the attempt is crucial, as each potential block represents a chance to directly win the point or radically disrupt the opponent's offense. In every important game, there is always one intense block which resets the momentum and this visualization shows the power of each successful block. Thus, blocking is not a passive defensive act but a high-risk, high-reward offensive weapon that shapes the psychological and tactical landscape of a match.")
)

```


## **Data Source Information**

Row {data-height=300}
-----------------------------------------------------------------------

### About the Data {#datasource}

```{r data-sc}
tags$div(
  style = "padding: 20px;",
  h3("📋 Data Sources"),
  tags$ul(
    tags$li(tags$strong("PlusLiga 2008-2025:"), 
            "Historical league data from Kaggle", 
            tags$a(href="https://www.kaggle.com/datasets/kacpergregorowicz/-mens-volleyball-plusliga-20082022", "compiled by KaKa", target="_blank"))
  ),
  hr(),
  h4("Research Question:"),
  p("What is the distribution for key statistics?"),
  hr(),
  p("Dashboard created with R flexdashboard • Updated: December 2025")
)
```
