---
title: "Maps (Filled) Guided Lab (Part 2)"
author: "Robert Thomas Kohn"
date: "10/28/25"
output: 
  pdf_document:
    fig_height: 5
    fig_width: 7
  html_document:
    fig_height: 5
    fig_width: 7
  word_document:
    fig_height: 5
    fig_width: 7
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
theme_set(theme_classic( ))
library(tidyverse)
library(lubridate)
library(mosaic)
library(Lock5Data)
library(Stat2Data)
library(dplyr)
library(ggrepel)
library(scales)
library(stringr)
library(usmap)
library(mapdata)
library(maps)
library(mapdata)
library(tibble)
library(gapminder)
library(viridis)
library(usdata)
```


# Filling Maps with Color

State abbreviations instead of state names: mutate( FullState = abbr2state( AbbState ) )
State has upper-case letters and we need lowercase: mutate( state = tolower( State ) )
Combine Both in one command: mutate( state = tolower( abbr2state( AbbState ) ) )

gf_polygon( Lat ~ Long , data = Dataset , group = ~ Group , fill = ~ Fill , color = " LineColor " , size = LineSize )
Conic for Entire US Map: coord_map("conic", lat0=30)
Equal Coords for Single State, Country, or World Map: coord_equal( ) or coord_fixed( )

# Graph 1: State map filled by numeric variable

```{r}

NewState77 = rownames_to_column( as.data.frame( state.x77 ) ) %>% # make row names a usable variable
rename(LifeExp = `Life Exp`, # Change messy variable names to clean ones
HSGrad = `HS Grad` ) %>% # Use ` (upper left on keyboard) instead of ‘ (right middle)
mutate(region = tolower( rowname ) ) # Make region variable with lowercase states
states <- map_data("state")
FullState77 <- full_join(NewState77, states, by = "region")
gf_polygon(lat ~ long, data = FullState77, group = ~group, fill = ~LifeExp, color = "white", size = 0.2)
gf_polygon(lat ~ long, data = FullState77, group = ~group, fill = ~LifeExp, color = "white", size = 0.2) %>% gf_refine(scale_fill_distiller(palette = "Spectral"))
gf_polygon(lat ~ long, data = FullState77, group = ~group, fill = ~LifeExp, color = "white", size = 0.2) %>% gf_refine(scale_fill_distiller(palette = "Spectral",direction = 1))
gf_polygon(lat ~ long, data = FullState77, group = ~group, fill = ~LifeExp, color = "white", size = 0.2) %>% gf_refine(scale_fill_distiller(palette = "Spectral",direction = 1, limits = c(67, 74),breaks = c(68, 70, 72, 74),name = "Life Expectancy\n(years)"))
gf_polygon(lat ~ long, data = FullState77, group = ~group, fill = ~LifeExp, color = "white", size = 0.2) %>% gf_refine(scale_fill_distiller(palette = "Spectral",direction = 1, limits = c(67, 74),breaks = c(68, 70, 72, 74),name = "Life Expectancy\n(years)")) %>% gf_refine(coord_map("conic", lat0 = 30)) %>%gf_labs(title = "Life Expectancy by State",subtitle = "Lower Life Expectancy Shown in Red",fill = "Life Expectancy")

```

# Graph 2: State map filled by categorical variable

```{r}

NewState77 = NewState77 %>%
mutate( binned_life = cut( LifeExp , breaks = 5 , # Create a categorical variable from a numeric one
labels = c( "Really Low" , "Low" , "Medium" , "High" , "Really High" ) ) )
names(NewState77)
head(NewState77)
FullState77 <- full_join(NewState77, states, by = "region")
gf_polygon(lat ~ long,data = FullState77, group = ~group,fill = ~binned_life, color = "white", size = 0.2) %>% gf_refine(scale_fill_brewer( palette = "Spectral",name = "Life Expectancy\nCategory" )) %>% gf_refine(coord_map("conic", lat0 = 30)) %>% gf_theme(theme_void()) %>% gf_labs(title = "Life Expectancy by State (Categorical)", subtitle = "Binned into 5 categories", fill = "Category")
gf_polygon(lat ~ long,data = FullState77, group = ~group,fill = ~binned_life, color = "white", size = 0.2) %>% gf_refine(scale_fill_brewer( palette = "Spectral", direction = "rev",name = "Life Expectancy\nCategory" )) %>% gf_refine(coord_map("conic", lat0 = 30)) %>% gf_theme(theme_void()) %>% gf_labs(title = "Life Expectancy by State (Categorical)", subtitle = "Binned into 5 categories", fill = "Category")#personally pref no reverse


```

# Graph 3: County map filled diverging sequence

```{r}

county <- map_data("county")
emp2 = viridis :: unemp %>% # Load unemp data from the viridis package
mutate( region = tolower( abbr2state( state ) ) ) # Make abbreviations state names instead
emp.county = full_join( emp2 , county , join_by( region , county == subregion ) )
gf_polygon(lat ~ long,data = emp.county,group = ~group,fill = ~rate, color = NA)
gf_polygon(lat ~ long,data = emp.county,group = ~group,fill = ~rate, color = NA) %>% gf_refine(scale_fill_gradient2(low = "blue",mid = "white", high = "red",midpoint = 5,name = "Unemployment\nRate (%)"))
gf_polygon(lat ~ long,data = emp.county,group = ~group,fill = ~rate, color = NA) %>% gf_refine(scale_fill_gradient2(low = "blue",mid = "white", high = "red",midpoint = 5,name = "Unemployment\nRate (%)")) %>%gf_refine(coord_map("conic", lat0 = 30)) %>%gf_labs(title = "Unemployment Rate by County",subtitle = "Blue = below 5%, Red = above 5%", fill = "Rate (%)")

```

# Graph 4: World map filled

```{r}

world <- map_data("world")
gapminder2007 <- gapminder %>%filter(year == 2007)
gapminder.newnames = full_join(gapminder2007 , Countries, join_by( country == gapminder ))
world.gapminder = left_join(world, gapminder.newnames , join_by( region == maps))
gf_polygon(lat ~ long,data = world.gapminder,group = ~group,fill = ~gdpPercap, color = "gray50", size = 0.1)
gf_polygon(lat ~ long,data = world.gapminder,group = ~group,fill = ~gdpPercap, color = "gray50", size = 0.1) %>%gf_refine(scale_fill_viridis_c(option = "plasma", name = "GDP per Capita\n(USD)",trans = "log10")) %>%gf_refine(coord_fixed(ratio = 1.3)) %>%gf_labs(title = "GDP per Capita by Country (2007)",subtitle = "Log scale transformation",fill = "GDP per Capita")


```