---
title: "Merging Datasets Open Ended Assignment"
author: "Robert Thomas Kohn"
date: "10/10/25"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
theme_set(theme_classic( ))
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyverse)
library(scales)
library(mosaic)
```


# Broadway 3 & 4 Merging

```{r}
Broadway3 <- read.csv("Broadway3.csv")
Broadway3_clean <- Broadway3 %>% select(Show, TonyNomination, StyleOfShow, CriticReview, TypeOfShow, NumberOfPerform)
Broadway4 <- read.csv("Broadway4.csv")
Broadway4_clean <- Broadway4 %>% select(NameofShow, TimeBway, ShowType, TonyNoms, Genre, Past, YearOpened) %>%rename(Show = NameofShow)
Broadway3_clean <- Broadway3_clean %>%mutate(Show = str_trim(Show))
Broadway4_clean <- Broadway4_clean %>%mutate(Show = str_trim(Show))
head(Broadway3_clean)
head(Broadway4_clean)
Broadway_combined <- Broadway3_clean %>% inner_join(Broadway4_clean, by = "Show")
head(Broadway_combined)
```



# Question: How do Tony nominations relate to show run time and genre?

I plan to make a scatter plot of TimeBway vs TonyNomination, colored by Genre

```{r}
gf_point(TimeBway ~ TonyNomination, data = Broadway_combined, color = ~Genre)

# Revised
Broadway_filtered <- Broadway_combined %>% filter(TimeBway <= 100)  # outliers

gf_point(TimeBway ~ TonyNomination, data = Broadway_filtered, color = ~Genre, size = 3, alpha = 0.7) %>%
  gf_abline(slope = 5, intercept = 0, color = "black", linetype = "dashed") %>% gf_labs(x = "Tony Nominations", y = "Months on Broadway", title = "More Nominations Increases Broadway Stay", subtitle = "Scatterplot of Tony Nominations vs Time on Broadway", caption = "Broadway3 & Broadway4(CSV)") %>% gf_theme(legend.position = "right")
```

As per usual, I dislike using the merging commands and sorting data but I managed to make relatively quick work of this. I was hoping to do something more extravagant but the majority of points got last due to not having matching columns. As such, I got a final graph with 3 total points which tells me nothing at all but I did refine it from the original one. I thought in the final it would be nice to remove outliers and add a general trend line. 