---
title: "Refinements Open Ended Assignment"
author: "Robert Thomas Kohn"
date: "9/15/25"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
theme_set(theme_bw())
```

# Refinement 1 - Begin Graphing in RStudio Open Ended Assignment 1

Question 1: Does a person having a child under 18 influence how often they fly annually?

```{r}
require(ggmosaic)
data(fly)
head(fly)
fly=na.omit(fly)


print("The Following is my Original Plot from Begin Graphing in RStudio Open Ended")

gf_bar( ~ has_child_under_18 , data = fly , fill = ~ flight_freq,position = position_dodge( ), title="Seperated Bar Chart | Explaining the Number of Flights with if they have a minor" )


print("The Following is my Revised Plot from Begin Graphing in RStudio Open Ended")
gf_bar( ~ has_child_under_18 , data = fly , fill = ~ flight_freq,position = position_dodge( ))%>%gf_labs(x = "Do they have a child?", y = "Count", title="Explaining Number of Flights if They Have a Child <18" , subtitle = "Seperated Bar Chart" , caption = "Dataset: fly.csv", fill="How often they fly")
```

I feel like my intial graph was very strong but I felt it was necessary to revise it because it had flight_freq which made it gross and the axes labels were very unappealing. Now I feel like it is significantly better but I still wonder if there is a better title that could be used. Its hard to incorporate it without it being too long. Perhaps if there is a way to wrap text but even then that may be too wordy and take away from the graph.

# Refinement 2 - Begin Graphing in RStudio Open Ended Assignment 2

Question 2: Is there a relationship between a State's Per Capita Income and the Winner of the 2008 Election?

```{r}
require(Stat2Data)
data(Election08)
head(Election08)
Election08$Winner <- ifelse(Election08$ObamaWin == 1, "Obama", "McCain")

print("The Following is my Original Plot from Begin Graphing in RStudio Open Ended")
gf_violin( Winner ~ Income , data = Election08 ,title="Violin Plot of Per Capita Income vs Winner of State")



print("The Following is my Revised Plot from Begin Graphing in RStudio Open Ended")
gf_violin( Winner ~ Income , data = Election08, fill = ~Winner)%>%gf_labs(x = "The Per Capita Income of the State", y = "Candidate", title="Violin Plot of Per Capita Income vs Winner of State" , subtitle = "The Results from the 2008 United States Election" , caption = "Dataset: Election08(Stat2Data)")

```

I feel like the difference between the two is night and day fully. Not only are the individual violins filled in with the proper party colors but this whole graph looks relatively professional. It shows the candidates along side the per capita income and clearly labeled key and title. I am very content with this one.



# The New Graphs

# Is there a relationship between the percent of College Graudates and  Democratic Lean in the 2016 Election?

```{r}
#numericvnum
require(Stat2Data)
data(Election16)
gf_point( Dem.Rep ~ BA , data = Election16 , alpha = 0.5 , title="Scatterplot Bachelors Percentage vs Democratic Lean")
gf_density_2d( Dem.Rep ~ BA , data = Election16, title="contour lines and each individual point") %>% gf_point(Dem.Rep ~ BA , data = Election16 , alpha = 0.5 )
gf_density_2d_filled( Dem.Rep ~ BA , data = Election16, title="contour lines and each individual point Filled") %>% gf_point(Dem.Rep ~ BA , data = Election16 , alpha = 0.5, color="white")

#REVISED
gf_point( Dem.Rep ~ BA , data = Election16 , alpha = 0.5)%>%gf_labs(x = "The Per Capita Income of the State", y = "Political Party Lean", title="Democrats are positively linked to Per Capita Income" , subtitle = "Results from the 2016 United States Election" , caption = "Dataset: Election16(Stat2Data)")%>%gf_hline(yintercept = 0, linetype = "solid", color = "black", size=1)%>%gf_text( 20 ~ 20 , hjust = 0, label = "Democrat Lean" , size = 6 , color = "blue" )%>%gf_text( -20 ~ 30 , hjust = 0, label = "Republican Lean" , size = 6 , color = "red" )
```

On this new graph, I started very bare bones but landed on this one as I felt like it illustrated an interesting design comparable to a win delta where its expected to be 0 for no relation. However, with this graph there was a clear relation so I did some touch ups. I added a horizontal line in to show the no relation line and added labels to show case what being above and below would mean. I also added labels and titles to make it accessible to the reader and giving a caption with the dataset and where it was accessed from. 

# Is there a relationship between the weather and if the dogs were walked?

```{r}
#CATEGORICAL
data(WalkTheDogs )
WalkTheDogs$Walked <- ifelse(WalkTheDogs$Walk == 1, "Yes", "No")
gf_bar( ~ Weather , data = WalkTheDogs  , fill = ~ Walked , position = position_fill( ),title="Position Filled in WalkTheDogs Bar Chart"  )
gf_bar(~Walked, data = WalkTheDogs , fill = ~Walked, show.legend = FALSE) %>% gf_facet_wrap(~Weather)
gf_bin2d(Walked~Weather, data=WalkTheDogs ) %>%
gf_refine(scale_fill_continuous(low = "white", high = "navy"))

#REVISED
gf_bar(~Walked, data = WalkTheDogs , fill = ~Walked, show.legend = FALSE) %>% gf_facet_wrap(~Weather, labeller = labeller(Weather = c("cold" = "Cold", "rain" = "Rain", "shine"="Shine")))%>%gf_labs(x = "The Weather and If the Dog was Walked", y = "How many occurences", title="Not Walking the Dog is Common in Any Weather" , subtitle = "Bar Graph Seperated by Weather" , caption = "Dataset: WalkTheDogs(Stat2Data)" )
```

I decided to go with this graph because I felt like it provided the most interested and compelling story. The first position filled one was certainly interesting but I felt like it told an incomplete story. I added labels with a good title to explain the story to those who were curious about how often the dog is walked.

# Is there a relationship between the rater's sex and how they rate someone's sexual dimorphism?

```{r}
#1n1
data(FaithfulFaces)
library(ggridges)
gf_boxplot( SexDimorph ~ RaterSex , data = FaithfulFaces , options,title="Boxplot")

gf_density(~SexDimorph, data = FaithfulFaces,title="Density") %>%
gf_facet_wrap(~RaterSex, ncol=1, strip.position = "right")

gf_density_ridges(RaterSex~SexDimorph, data=FaithfulFaces, fill=~RaterSex, show.legend = FALSE, scale = 2, alpha = 0.5,title="Density Ridges Plot") 

#

#REVISED
gf_density(~SexDimorph, data = FaithfulFaces, fill=~RaterSex) %>% gf_facet_wrap(~RaterSex, ncol=1, strip.position = "right", labeller = labeller(RaterSex = c("F" = "Female", "M" = "Male"))) %>%gf_labs(x = "Rating of Sexual Dimorphism", y = "Frequency of Values", title="Women Rate Higher Sexual Dimorphism than Men" , subtitle = "Density of Sexual Dimorphism Seperated by Sex" , caption = "Dataset: FaithfulFaces(Stat2Data)") + scale_fill_manual(values = c("F" = "lightpink", "M" = "lightblue"), labels = c("F" = "Female", "M" = "Male"), name = "Sex of Rater")
```

This density plot was the most intriguing when I did this as it gave me a unique way to look at how sexual dimorphism is perceived and the proportion at each part. This graph was very fun to edit and add on to. Obviously it was labelled on the axes and titles as with the others. I also labeled the key and gave colors to each density plot making it look nice. I was very happy with the way it turned out and it was worth the struggles with the code. 