---
title: "Computational Probability & Monte Carlo Simulations"
author: "Robert Kohn"
date: "Fall 2025"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(mosaic)
library(tidyverse)
library(mnormt)
library(DescTools)
library(cubature)
library(ggplot2)
library(scales)
set.seed(67)
```

# Introduction & Mathematical Framework

Monte Carlo simulation provides a computational bridge between theoretical mathematical statistics and empirical probability. As a Mathematics and Statistics double major, this project investigates how computational algorithms verify theoretical expectations, asymptotic variance convergence, joint bivariate distributions, and multi-dimensional integration.

# Part 1: Discrete Distributions & Law of Large Numbers

## Binomial Distribution Moment Convergence

Consider a sequence of $n = 100$ independent Bernoulli trials with success probability $p = 0.30$:
$$X \sim \text{Binomial}(n = 100, p = 0.30)$$

Theoretical moments:
$$E[X] = np = 100 \times 0.30 = 30$$
$$\text{Var}(X) = np(1 - p) = 100 \times 0.30 \times 0.70 = 21.00$$

```{r binom-sim}
n_draws <- 100000
n_trials <- 100
p_binom <- 0.3

binom_sim <- rbinom(n_draws, size = n_trials, prob = p_binom)
binom_var_mc <- var(binom_sim)
binom_var_th <- n_trials * p_binom * (1 - p_binom)

cat(sprintf("Monte Carlo Variance: %.4f | Theoretical: %.4f | Absolute Error: %.6f\n",
            binom_var_mc, binom_var_th, abs(binom_var_mc - binom_var_th)))
```

The running sample variance demonstrates asymptotic convergence toward the theoretical variance of 21.0:

```{r binom-plot, echo=FALSE, fig.width=8, fig.height=4.5}
sample_steps <- seq(500, n_draws, by = 1000)
running_var <- sapply(sample_steps, function(k) var(binom_sim[1:k]))
conv_df <- data.frame(Sample_Size = sample_steps, Running_Variance = running_var)

ggplot(conv_df, aes(x = Sample_Size, y = Running_Variance)) +
  geom_line(color = "#1f77b4", linewidth = 1) +
  geom_hline(yintercept = binom_var_th, color = "#d62728", linetype = "dashed", linewidth = 1) +
  scale_x_continuous(labels = comma) +
  labs(
    title = "Monte Carlo Variance Convergence: Binomial(100, 0.3)",
    x = "Sample Size (N)",
    y = "Sample Variance"
  ) +
  theme_minimal()
```

## Linear Combinations of Independent Discrete Variables

Let $X \sim \text{Poisson}(\lambda = 2)$ and $Y \sim \text{Geometric}(p = 0.50)$ (number of failures before first success). By linearity of expectation:
$$E[X + Y] = E[X] + E[Y] = \lambda + \frac{1 - p}{p} = 2 + \frac{0.50}{0.50} = 3.00$$

```{r sum-sim}
x_sim <- rpois(n_draws, lambda = 2)
y_sim <- rgeom(n_draws, prob = 0.5)
sum_sim <- x_sim + y_sim

cat(sprintf("Monte Carlo E[X + Y]: %.4f | Theoretical: %.4f | Error: %.6f\n",
            mean(sum_sim), 3.0, abs(mean(sum_sim) - 3.0)))
```

## Negative Binomial Tail Probabilities

For $X \sim \text{NegBin}(r = 5, p = 0.65)$, we calculate the tail probability $P(X > 6)$:
$$P(X > 6) = 1 - \sum_{k=0}^6 \binom{k + r - 1}{k} (1 - p)^k p^r$$

```{r nb-tail}
nb_sim <- rnbinom(n_draws, size = 5, prob = 0.65)
p_gt_6_mc <- mean(nb_sim > 6)
p_gt_6_th <- 1 - sum(sapply(0:6, function(k) choose(k + 5 - 1, k) * (1 - 0.65)^k * (0.65)^5))

cat(sprintf("Monte Carlo P(X > 6): %.5f | Theoretical: %.5f | Error: %.6f\n",
            p_gt_6_mc, p_gt_6_th, abs(p_gt_6_mc - p_gt_6_th)))
```

# Part 2: Continuous Distributions & Monte Carlo Integration

Simulating across continuous parametric families: Uniform $\mathcal{U}(3, 10)$, Exponential $\text{Exp}(\lambda = 2)$, Chi-squared $\chi^2(5)$, and Beta $\text{Beta}(2, 4)$:

```{r cont-dist, fig.width=8.5, fig.height=5}
unif_draws  <- runif(10000, min = 3, max = 10)
exp_draws   <- rexp(10000, rate = 2)
chisq_draws <- rchisq(10000, df = 5)
beta_draws  <- rbeta(10000, shape1 = 2, shape2 = 4)

dist_df <- data.frame(
  Value = c(unif_draws, exp_draws, chisq_draws, beta_draws),
  Distribution = factor(rep(c("Uniform(3, 10)", "Exponential(rate=2)", "Chi-Squared(df=5)", "Beta(2, 4)"),
                            each = 10000))
)

ggplot(dist_df, aes(x = Value, fill = Distribution)) +
  geom_density(alpha = 0.6, color = "black", linewidth = 0.3) +
  facet_wrap(~ Distribution, scales = "free", nrow = 2) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Empirical Density Functions of Continuous Probability Distributions", x = "x", y = "Density") +
  theme_minimal() +
  theme(legend.position = "none")
```

# Part 3: Bivariate Normal Distribution & Hypothesis Testing

Let $\mathbf{X} = (X_1, X_2)^T \sim \mathcal{N}_2(\boldsymbol{0}, \mathbf{I}_2)$. Under independence, the joint density factors:
$$f(x_1, x_2) = f_{X_1}(x_1) f_{X_2}(x_2)$$

```{r biv-norm}
mu_vec <- c(0, 0)
cov_mat <- diag(2)

joint_origin <- dmnorm(c(0, 0), mean = mu_vec, varcov = cov_mat)
prod_marg    <- dnorm(0) * dnorm(0)
cat(sprintf("Joint PDF at (0, 0): %.6f | Product of Marginals: %.6f\n", joint_origin, prod_marg))

# Cumulative probability over rectangular region [-2, 2] x [-2, 2]
p_rect_joint <- pmnorm(c(2, 2), mean = mu_vec, varcov = cov_mat) -
                pmnorm(c(-2, 2), mean = mu_vec, varcov = cov_mat) -
                pmnorm(c(2, -2), mean = mu_vec, varcov = cov_mat) +
                pmnorm(c(-2, -2), mean = mu_vec, varcov = cov_mat)
p_rect_marg  <- (pnorm(2) - pnorm(-2))^2
cat(sprintf("Joint CDF Rectangle: %.6f | Marginal Product: %.6f\n", p_rect_joint, p_rect_marg))
```

## Sample Correlation & Hypothesis Testing

Simulating $N = 2,000$ points from standard bivariate normal:
$$H_0: \rho = 0 \quad \text{vs.} \quad H_1: \rho \neq 0$$

```{r cor-test}
biv_sim <- rmnorm(2000, mean = mu_vec, varcov = cov_mat)
cor_res <- cor.test(biv_sim[, 1], biv_sim[, 2])
cat(sprintf("Sample Pearson r: %.4f | p-value: %.4f | Decision: %s\n",
            cor_res$estimate, cor_res$p.value,
            ifelse(cor_res$p.value > 0.05, "Fail to reject H0 (Consistent with independence)", "Reject H0")))
```

```{r biv-plot, echo=FALSE, fig.width=6.5, fig.height=6}
biv_df <- data.frame(X1 = biv_sim[, 1], X2 = biv_sim[, 2])
ggplot(biv_df, aes(x = X1, y = X2)) +
  geom_point(alpha = 0.4, color = "#2ca02c", size = 1.5) +
  geom_density_2d(color = "black", linewidth = 0.5) +
  coord_fixed(xlim = c(-3.5, 3.5), ylim = c(-3.5, 3.5)) +
  labs(
    title = "Simulated Bivariate Standard Normal Distribution (N = 2,000)",
    subtitle = sprintf("Zero correlation structure: r = %.4f (p-value = %.3f)", cor_res$estimate, cor_res$p.value),
    x = "X1", y = "X2"
  ) +
  theme_minimal()
```

# Part 4: Numerical Quadrature & Combinatorics

## 1D and Multidimensional Integration

Integrating $g(x) = x^2 + 3x - 4$ over $[0, 2]$:
$$\int_0^2 (x^2 + 3x - 4) dx = \left[ \frac{x^3}{3} + \frac{3x^2}{2} - 4x \right]_0^2 = \frac{8}{3} + 6 - 8 = \frac{2}{3} \approx 0.666667$$

```{r num-int}
quad_1d <- integrate(function(x) x^2 + 3*x - 4, lower = 0, upper = 2)
cat(sprintf("Numerical Quadrature: %.6f | True Value: 0.666667\n", quad_1d$value))

# 2D Adaptive Integration over [0, 1] x [0, 1.5]
poly_2d <- function(vec) vec[1]^2 + 3*vec[1]*vec[2] - 4*vec[2]^3
quad_2d <- adaptIntegrate(poly_2d, lowerLimit = c(0, 0), upperLimit = c(1, 1.5))
cat(sprintf("2D Adaptive Quadrature: %.6f | Error estimate: %.2e\n", quad_2d$integral, quad_2d$error))
```

## Combinatorics

Enumerating all subsets of size $k = 3$ chosen from $n = 8$ items without replacement:
$$\binom{8}{3} = \frac{8 \times 7 \times 6}{3 \times 2 \times 1} = 56$$

```{r comb-set}
flavors <- c("Vanilla", "Chocolate", "Strawberry", "Cookies & Cream",
             "Cookie Dough", "Mint Chip", "Butter Pecan", "Coffee")
combos <- CombSet(flavors, 3, repl = FALSE, ord = FALSE)
cat(sprintf("Empirical combinations generated: %d | choose(8, 3): %d\n", nrow(combos), choose(8, 3)))
```
